\documentclass[twocolumn]{article}

% Essential packages
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage[authoryear,round]{natbib}
\usepackage{appendix}


% Page geometry
\usepackage[letterpaper,
            includeheadfoot,
            layoutsize={8.125in,10.875in},
            layouthoffset=0.1875in,
            layoutvoffset=0.0625in,
            left=38.5pt,
            right=43pt,
            top=43pt,
            bottom=44pt,
            headheight=0pt,
            headsep=-2pt,
            columnsep=0.3125in,
            footskip=25pt,
            marginparwidth=38pt]{geometry}

% Authors and affiliations
\usepackage{authblk}

% Float control
\renewcommand{\topfraction}{0.9}
\renewcommand{\bottomfraction}{0.9}
\renewcommand{\textfraction}{0.05}
\renewcommand{\floatpagefraction}{0.9}

% Math operators
\DeclareMathOperator*{\mean}{mean}
\DeclareMathOperator*{\std}{std}
\DeclareMathOperator*{\AbsDiff}{AbsDiff}

% Bibliography settings
\setlength{\bibsep}{0pt plus 0.3ex}

% First page footer
\usepackage{fancyhdr}
\fancypagestyle{firstpage}{%
  \fancyhf{}%
  \fancyfoot[C]{\footnotesize
    This version has not been peer reviewed. Comments and feedback are welcome.}%
  \renewcommand{\headrulewidth}{0pt}%
}

% Caption formatting
\usepackage{caption}
\captionsetup{labelfont=bf, font=small}

% Author/affiliation fonts
\renewcommand\Authfont{\large}
\renewcommand\Affilfont{\normalsize}

% Autoref customization
\renewcommand{\tableautorefname}{Tab.}
\renewcommand{\figureautorefname}{Fig.}
\renewcommand{\equationautorefname}{Eq.}
\renewcommand{\appendixautorefname}{Appx.}

% TODO: confirm Jorge's affiliation number and institution
\title{\LARGE Probing Warmth and Competence Representations in LLM Hiring Decisions}
\date{\today}
\author[1]{Emrecan Ulu}
\author[2]{Jorge Lastra Cerda}
\affil[1]{University of Konstanz}
\affil[2]{University of Konstanz}  % TODO: update if different

\begin{document}
\twocolumn[
  \begin{@twocolumnfalse}
\maketitle
\thispagestyle{firstpage}


\begin{abstract}
\noindent
% TODO: write abstract once results are finalized.
% Suggested structure (3--4 sentences):
% (1) Discrimination in hiring is well-documented in humans and increasingly in AI.
% (2) We probe whether warmth and competence (Stereotype Content Model) are
%     encoded as linear directions in the residual stream of Gemma 3 12B.
% (3) We steer those directions and measure causal shifts in callback recommendations,
%     benchmarked against Gallo and Hausladen (2024).
% (4) Findings and implications.
\texttt{Due: Week 8}
\vspace{0.50cm}
\end{abstract}
  \end{@twocolumnfalse}
]


\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\footnotetext[0]{Corresponding author: TODO@uni-konstanz.de}  % TODO: update email
\renewcommand{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

% ─────────────────────────────────────────────────────────────────────────────
\section*{Introduction}
% ─────────────────────────────────────────────────────────────────────────────

Discrimination in hiring is one of the most persistent and consequential
forms of social inequality.
Access to work is not a neutral outcome: it shapes social mobility,
material security, and the possibilities available to people across their
lives.
One of the most cited field experiments documenting the scope of this
problem is \citet{bertrand2004emily}, who sent nearly 5{,}000 matched
resumes to employers in Boston and Chicago, varying only the perceived
race of the applicant through the name on the resume.
Resumes assigned White-sounding names received approximately 50~percent
more callbacks than otherwise identical resumes assigned
African-American-sounding names, a gap that held across industries,
occupations, and firm sizes.
Similar audits have since documented comparable patterns for gender, age,
disability, religion, sexual orientation, and national origin,
establishing that differential treatment in hiring is not an isolated
phenomenon but a structural feature of labor markets in North America and
beyond \citep{correll2007getting, neumark2019older, tilcsik2011pride}.

Social psychology offers a theoretical framework for understanding why
these disparities exist and what drives their magnitude.
The Stereotype Content Model \citep[SCM;][]{fiske2002model} proposes that
social perception is organized along two fundamental dimensions: warmth,
which captures perceived intentions and trustworthiness, and competence,
which captures perceived ability and effectiveness.
Groups stereotyped as low on both dimensions elicit contempt and active
harm, while groups stereotyped as high on both elicit admiration and
active facilitation \citep{cuddy2007biasmap}.
\citet{gallo2024warmth} tested this model directly against hiring outcomes.
In a meta-analysis pooling callback data from over two decades of North
American correspondence studies, they showed that human warmth and
competence ratings of social signals predict callback disparities with
substantial accuracy.
The way people perceive the warmth and competence of a social group is
not just a psychological description: it is a working predictor of
whether members of that group get called back for a job interview.

Large language models are now entering this picture.
They are being used to screen resumes, rank candidates, and support or
automate hiring decisions at scale.
This raises an urgent question: do these systems reproduce the
discriminatory patterns documented in human hiring, and if so, what is
the underlying mechanism?
Behavioral audit studies have already begun to answer the first part.
\citet{an2024llm} found systematic disparities in GPT-3.5 callback
recommendations across racial and ethnic groups.
\citet{gaebler2024auditing} audited multiple large language models and found
persistent race and gender disparities with potential legal implications
under equal employment opportunity law.
What these studies cannot tell us is where in the model this differential
treatment originates, and how it could be addressed at the source rather
than patched at the output.

% ─────────────────────────────────────────────────────────────────────────────
\section*{Literature}
% ─────────────────────────────────────────────────────────────────────────────

Four bodies of literature converge on the problem this paper addresses,
and each leaves a specific gap.

\subsection*{Correspondence Studies}

Correspondence studies have proven that discrimination in hiring is real
and measurable across race, gender, age, disability, and national origin.
What they cannot tell us is why it happens at the level of cognitive
mechanism: they describe the outcome but not the process that produces it.

\subsection*{The Stereotype Content Model}

The SCM answers that question for humans.
Warmth and competence are the two dimensions through which social groups
are perceived, and those perceptions predict who gets called back, as
\citet{gallo2024warmth} demonstrated directly.
But the SCM was built entirely within human psychology.
Nobody has tested whether this same structure exists inside a machine.
If a language model trained on human-generated text has internalized
warmth and competence as internal representations, that would be a
meaningful finding; if it has not, that is equally important to know.

\subsection*{Behavioral Audits of Large Language Models}

Behavioral audits of large language models have confirmed that AI systems
also discriminate in hiring \citep{an2024llm, gaebler2024auditing}.
But by treating the model as a black box, these studies can document that
disparities exist without explaining where in the model they originate or
what internal structure produces them.
Knowing that a model recommends against an applicant with a Black-sounding
name more often than an identical applicant with a White-sounding name is
useful, but it does not tell us how to fix it or where the bias lives.

\subsection*{Mechanistic Interpretability}

Mechanistic interpretability gives us tools to open that box.
Representation engineering \citep{zou2023repeng} and activation addition
\citep{turner2023activation} have shown that high-level concepts are encoded as
linear directions in large language model residual streams and can be
causally manipulated.
Most recently, \citet{sofroniew2026emotion} applied this framework to emotion
concepts in Claude Sonnet, showing that internal emotion representations
causally influence model behavior.
But these tools have never been applied to warmth and competence as social
stereotype dimensions, and never connected to hiring discrimination.

\subsection*{This Paper}

This paper sits at the intersection of all four gaps and makes a
methodological contribution by connecting the tools of mechanistic
interpretability to the empirical tradition of correspondence studies.
We take the conceptual framework from the SCM, the empirical benchmark
from correspondence studies, the motivating question from LLM bias audits,
and the methodological tools from mechanistic interpretability.
We extract and validate warmth and competence direction vectors from three
open-weights models spanning three independent architectural lineages, test
whether those representations are causally functional, and connect the
resulting callback disparities to the human benchmark of
\citet{gallo2024warmth}.
Doing so allows us to move the study of LLM hiring bias from behavioral
description to mechanistic account: not only documenting that the disparity
exists, but locating where in the model it may originate and what
representational structure underlies it.

% ─────────────────────────────────────────────────────────────────────────────
\section*{Methods}
% ─────────────────────────────────────────────────────────────────────────────

\paragraph{Story Corpus.}
To construct concept direction vectors independently of demographic identity,
we generated 200 short narratives (approximately 90--150 words each) distributed
evenly across four conditions: high warmth, low warmth, high competence, and
low competence (50 stories per condition).
Stories were written around 50 distinct everyday topics spanning ten domains
(workplace, learning, community, sport, travel, and others), with each topic
appearing once per condition so that situational vocabulary is balanced across
the high and low poles.
All stories feature a nameless, gender-neutral third-person protagonist (``they'').
This design choice was motivated by pilot work showing that AI-generated
concept stories exhibit systematic demographic skew: low-warmth and
low-competence conditions were overwhelmingly assigned male, Anglo-named
protagonists, while high conditions were assigned female and demographically
diverse protagonists.
Without neutralisation, direction vectors would encode demographic expectations
alongside the target social dimension, confounding the probe with the very bias
we seek to study.
By removing identity signals entirely, each story's activation is driven by
depicted behaviour rather than group membership.
Stories were generated by a large language model and validated by automated
checks for forbidden-vocabulary violations, length balance across conditions,
and demographic neutrality.
The use of AI-generated stimuli raises concerns about circularity that we address
in the limitations.
The current corpus of 200 stories is a pilot; we plan to expand to 400 stories
as data collection continues.

\paragraph{Activation Extraction.}
We used TransformerLens \citep{nanda2022transformerlens} to extract
residual-stream activations from three open-weights instruction-tuned models:
Gemma-3-12B-it \citep{gemmateam2025gemma3} as the primary model, and
Qwen3-14B \citep{qwen2025qwen3} and Llama-3.1-8B-Instruct
\citep{grattafiori2024llama3} as cross-architecture replications.
For each model we captured the post-layer residual-stream state at the layer
corresponding to approximately 66\% of model depth (layer~31 of~48 for
Gemma-3-12B; layer~26 of~40 for Qwen3-14B; layer~20 of~32 for Llama-3.1-8B).
For each story, activations were mean-pooled across token positions from
position~50 onward, skipping the prompt prefix, yielding one
$d_\mathrm{model}$-dimensional vector per story.
Warmth and competence direction vectors were then constructed as mean
differences:
$v_W = \bar{x}_\mathrm{high\,warmth} - \bar{x}_\mathrm{low\,warmth}$
and
$v_C = \bar{x}_\mathrm{high\,comp.} - \bar{x}_\mathrm{low\,comp.}$
All three models were run under strictly identical conditions
(same stories, same pooling rule, same seed 20260527) to enable
cross-architecture comparison.

\paragraph{PCA Denoising.}
The warmth and competence direction vectors share a substantial common
component reflecting general narrative valence: stories portraying protagonists
positively (whether warmly or competently) occupy a similar region of the
residual stream, as do negatively valenced stories.
To separate conceptual content from evaluative polarity, we follow the
neutral-corpus denoising procedure of \citet{sofroniew2026emotion}.
We construct a neutral corpus of 1,500 Wikipedia introduction paragraphs,
length-matched to the concept stories (90--200 words) and filtered to exclude
emotionally charged or violent content.
Residual-stream activations are extracted at the same layer and pooling rule.
Principal component analysis (PCA) is fitted on these neutral activations;
the top components collectively explaining at least 50\% of neutral variance
are then projected out of both direction vectors.
The resulting denoised vectors are used for all causal steering experiments.
% TODO: report pre/post cosine(warmth, competence) once denoising job completes.

\paragraph{Probe Validation.}
We assess the validity of each direction vector using five complementary
metrics.
(i)~\textit{5-fold cross-validated accuracy}: a logistic regression classifier
is trained on full residual-stream representations within each target contrast;
accuracy is averaged across five stratified folds.
(ii)~\textit{Topic-holdout accuracy}: a stricter variant using GroupKFold,
where all stories from a given topic are assigned to the same fold, testing
whether the probe generalises to entirely unseen situations rather than merely
to unseen phrasings of familiar topics.
(iii)~\textit{Cohen's $d$}: the standardised mean difference between
high- and low-condition projections onto the direction vector, compared against
an empirical null distribution of 1,000 random unit vectors.
(iv)~\textit{Split-half cosine stability}: each condition's 50 stories are
randomly split into halves of 25; a direction is computed from each half and
the cosine between them is reported.
(v)~\textit{Cross-axis classification}: using fold-local standardised
projections, we test whether the warmth direction predicts competence labels
and vice versa, quantifying the degree of shared evaluative content.
Per-story Spearman rank correlations across model pairs are also computed to
test whether the three models converge on a shared cross-architecture construct.

\paragraph{Causal Steering.}
To test whether the extracted directions causally influence model behaviour
rather than merely correlating with it, we apply activation steering
\citep{turner2023activation} via TransformerLens hooks at the probe layer.
Steering strength is expressed relative to the mean residual-stream norm at
that layer, ensuring comparability across models whose activation scales
differ by orders of magnitude.
We use a local regime of $\alpha \in \{-0.10,\,-0.05,\,0,\,+0.05,\,+0.10\}$
times the mean norm; a broader range is used as a saturation diagnostic.
For \textit{concept-level} evaluation, the model answers ``Is the protagonist
high in warmth/competence? Answer only Yes or No'' on held-out topics (40
training / 10 test, determined by topic index).
The outcome is the next-token logit margin $\Delta = \mathrm{logit}(\text{Yes}) -
\mathrm{logit}(\text{No})$.
For \textit{hiring-level} evaluation, the steering hooks are applied while
the model processes resume-evaluation prompts, and the outcome is the
callback logit margin.
We test six directions: the dense target direction, a decoded
Gemma Scope~2 SAE direction \citep{deepmind2025gemmascope2}, an axis-specific
component, a shared component, the opposing concept axis, and a random
orthogonal control.
Uncertainty is estimated via paired topic bootstrap resampling.
% TODO: report hiring-level results once the evaluation job completes.

\paragraph{Benchmark.}
We compare model callback disparities against the warmth and competence
ratings and callback rates from \citet{gallo2024warmth}, which provides
empirical grounding in human correspondence studies.

\paragraph{Data.}
Hiring stimuli are built around the 282 applicant names and category-level
signals from \citet{gallo2024warmth}, paired with a fixed administrative
assistant resume template to isolate the effect of the social signal.

% ─────────────────────────────────────────────────────────────────────────────
\section*{Results}
% ─────────────────────────────────────────────────────────────────────────────

\paragraph{Warmth and Competence Are Linearly Encoded Across Model Families.}
A cross-validated logistic regression probe achieves 100\% accuracy on
both the warmth and competence contrasts across all three tested models,
in both standard 5-fold cross-validation and the stricter topic-holdout
variant where entire situational topics are withheld from training.
The perfect topic-holdout accuracy indicates that the probe captures
something that generalises across entirely new social situations, not
merely unseen phrasings of familiar topics.
The separation strengths, measured as Cohen's~$d$ relative to a null
distribution of 1,000 random unit directions, are exceptional for all
models: for Gemma-3-12B ($d = 2.70$/$2.83$ for warmth/competence,
$z = 3.9$/$3.7$), Qwen3-14B ($d = 8.97$/$9.97$, $z = 14.1$/$14.6$),
and Llama-3.1-8B ($d = 8.48$/$9.07$, $z = 15.0$/$15.1$), zero of 1,000
random directions exceeded the target direction in any model.
The split-half cosine stability of the Gemma direction vectors (0.83 for
warmth, 0.88 for competence) confirms that these directions are stable
properties of the conditions rather than artefacts of specific stories.
Critically, the three models do not merely separate high from low
conditions in isolation: per-story Spearman rank correlations across model
pairs reveal that they agree substantially on which specific stories are
warm or cold ($\rho = 0.76$--$0.98$ for warmth; $0.78$--$0.99$ for
competence), indicating convergence on a shared cross-architecture
construct rather than independent idiosyncratic representations.
The cosine similarity between warmth and competence direction vectors
(0.749 for Gemma; 0.51--0.54 for Qwen3 and Llama) and the cross-axis
classification accuracy (0.82--0.87 for Gemma; $\approx$1.00 for
Qwen3/Llama) indicate that the two directions share a common evaluative
component reflecting the positive-vs-negative framing of the stories.
This result is addressed by the PCA denoising step described above;
denoised results are forthcoming.
% TODO: update with denoised vector metrics once PCA job completes.

\paragraph{Warmth and Competence Directions Are Causally Functional.}
To move beyond representational description, we applied activation steering
to the warmth and competence direction vectors in Gemma-3-12B and
Gemma-3-27B on held-out concept judgement tasks.
Before intervention, the model answered direct warmth/competence Yes/No
questions with 100\% accuracy on held-out topics, with baseline high--low
logit-margin gaps of 16.00/14.18 (12B) and 25.93/23.64 (27B).
Steering the dense target directions in the local regime produced positive,
approximately linear causal effects in all four model-by-axis combinations
(slopes: 12B warmth 27.79, $R^2 = 0.956$; 12B competence 15.11,
$R^2 = 0.915$; 27B warmth 12.89, $R^2 = 0.990$; 27B competence 8.85,
$R^2 = 0.826$).
The steering effect does not support a clean two-module interpretation,
however: the opposing concept axis also shifts judgements substantially in
several conditions.
Rather than treating this as a failure of specificity, we read it as a
positive finding that connects to the structure of the human benchmark.
\citet{gallo2024warmth} observe that warmth and competence ratings reduce
to a dominant shared principal component that predicts callback rates;
our steering results suggest that the same evaluative structure is
represented inside the model, such that moving along either dimension
also moves the shared positive-vs-negative appraisal.
Cross-scale feature-profile correlations between matched Gemma-3-12B and
Gemma-3-27B Gemma Scope~2 features (mean $r = 0.49$--$0.66$, exceeding
a 500-permutation row-shuffle null at $p = .002$ for all five vector
types) further indicate that the representational organisation is conserved
across scale.

\paragraph{The Model's Internal Probe Scores Track Human Social Perceptions.}
Before testing causal effects on hiring outcomes, we establish that the
model's internal probe scores reflect human social perception.
For each of the 282 applicant names rated by \citet{gallo2024warmth},
we extract the model's residual-stream activation at the probe layer and
project it onto the warmth and competence direction vectors.
These model-internal scores correlate positively with human warmth and
competence ratings: Spearman $\rho = 0.355$ for warmth and $\rho = 0.230$
for competence across all 282 names.
Human warmth and competence ratings also predict the model's baseline
callback margin ($\rho = +0.242$, $p < .001$; $\rho = +0.201$, $p < .01$),
while the model's own probe scores predict callback margin in the same
direction, though at lower magnitude
($\rho = +0.139$, $p = .084$; $\rho = +0.145$, $p = .067$).
These results indicate that names representing groups perceived as
warmer or more competent by humans are also represented as such
internally by Gemma-3-12B, and that this internal representation aligns
directionally with the model's callback inclination.
At Gemma-3-27B the probe-human correlations are slightly stronger
($\rho = 0.381$ for warmth; $\rho = 0.283$ for competence), suggesting
that the larger model encodes human social perceptions with marginally
higher fidelity.
However, the relationship between internal probe scores and the callback
margin reverses at 27B: names the model represents as warmer or more
competent internally receive \emph{lower} callback margins
($\rho = -0.160$ and $-0.156$ respectively), opposite to the 12B
pattern.
This reversal coincides with a dramatically different baseline response
regime: the 27B model recommends callbacks for an implied
$P(\text{Yes}) = 0.760$ of applicants, compared with $0.452$ at 12B,
and its callback margins are substantially more variable
($\mathit{SD} = 0.411$ vs $0.140$).
Together these results indicate that scale improves the fidelity with
which the model's internal axes track human perceptions, but shifts the
decision-making regime in a way that decouples those axes from the
direction of the callback signal.

\paragraph{Steering Warmth and Competence Causally Shifts Callback Recommendations.}
To test whether directly manipulating the warmth and competence
representations moves the model's hiring callback decision, we applied
activation steering during resume evaluation across 60 applicant names
drawn from the \citet{gallo2024warmth} rated name set.
The baseline callback margin was $-0.194$ ($\mathit{SD} = 0.130$),
corresponding to an implied $P(\text{Yes}) = 0.452$ — the model is
initially slightly disinclined to recommend a callback for the average
applicant on the standardised administrative-assistant profile.
Steering the warmth direction produced a positive, approximately linear
causal effect on the callback margin (slope $= +12.95$ $\Delta$margin
per unit of mean residual norm, $R^2 = 0.924$): increasing warmth
raised the model's callback inclination and decreasing warmth
reduced it monotonically across all tested strengths.
The competence direction also produced a positive mean slope
(slope $= +9.06$, $R^2 = 0.663$), but the effect was non-monotonic:
at the $-0.25$ strength level the mean delta was $+3.90$ rather than
the expected negative value, suggesting that moderate downward
competence steering does not reliably suppress the callback margin at
this steering resolution.
The warmth causal effect at 12B is therefore more robust and linear
than the competence effect, and constitutes the cleaner bridge from
internal representation to hiring decision at this scale.
Replicating the causal evaluation at Gemma-3-27B yields a qualitatively
different pattern.
The warmth steering effect is essentially absent (slope $= +1.09$,
$R^2 = 0.026$) and non-monotonic: decreasing warmth at $\alpha = -0.50$
raises rather than lowers the callback margin on average, while
increasing warmth at $\alpha = +0.25$ and $+0.50$ produces marginal
or negative effects.
The competence direction shows a small positive linear trend
(slope $= +2.88$, $R^2 = 0.340$), but positive steering at
$\alpha = +0.25$ and $+0.50$ suppresses rather than raises the
callback margin.
At 27B the concept directions that cleanly controlled explicit
warmth/competence judgements in the concept-level causal evaluation
do not translate into equivalent control over hiring decisions,
indicating a dissociation between representational control and
decision-level causality that depends on model scale.

\paragraph{Model Callback Disparities Mirror the Human Benchmark.}
% TODO: wire in demographic grouping (race, gender, national origin)
% from Gallo--Hausladen category files and join human callback rates
% from the meta-analysis data. Report model callback disparity per group
% vs human callback disparity, and run mediation (name ->
% model warmth/competence probe -> callback margin).
% Add Gemma-3-27B comparison for scale robustness once available.

% ─────────────────────────────────────────────────────────────────────────────
\section*{Discussion and Conclusion}
% ─────────────────────────────────────────────────────────────────────────────
% TODO: expand once hiring-level results are available.

That large language models trained on human-generated text would encode
warmth and competence as internally recoverable representations is,
in one sense, not surprising: these models distil patterns from the
same texts in which warmth and competence shape social perception.
What the present results establish empirically is that this encoding is
linear, geometrically stable, causally functional, and shared across
architecturally distinct models from independent laboratories.
Each of these properties matters: linearity makes the representation
accessible to lightweight probes and steering interventions; causal
functionality means the representation is not merely an epiphenomenon
but participates in the model's judgements; and cross-architecture
convergence suggests the finding reflects a property of training data
and the objectives of language modelling rather than a quirk of any
single design.
Together, the results suggest that LLMs have internalised a structure
of social perception analogous to the SCM, not through any explicit
design choice, but as a consequence of learning from text produced by
humans who perceive the world through it.
This has concrete implications.
It points to a mechanism for the callback disparities documented in
behavioral audits: differential treatment in hiring may arise not
merely from surface-level name associations but from how the model
represents the social meaning those names carry.
The scale comparison between Gemma-3-12B and Gemma-3-27B complicates
this picture in an instructive way.
The 27B model encodes warmth and competence with marginally higher
fidelity relative to human ratings, yet its hiring decisions are not
more discriminatory; on the contrary, it operates in a higher-baseline
regime ($P(\text{Yes}) = 0.76$) in which the warmth/competence probe
scores are negatively, not positively, associated with callback
inclinations, and in which activation steering of the concept directions
fails to produce systematic causal effects.
This dissociation --- better representation, weaker causal bridge to
decisions --- suggests that model capability and hiring bias do not
scale together simply.
A larger model may have internalised the SCM structure more precisely
while simultaneously developing decision processes that are less
directly governed by those representations, possibly reflecting the
influence of instruction tuning and reinforcement learning from human
feedback on the decision layer.
This finding points to a methodological implication for bias auditing:
probing internal representations and measuring output disparities are
complementary but not interchangeable, and a model that scores well on
one measure may not score well on the other.
It also opens a path to addressing the problem at the source --- by
intervening on the representation rather than filtering outputs after
the fact --- and provides a methodological template for connecting the
computational tools of mechanistic interpretability to the empirical
tradition of correspondence studies in the social sciences.

\paragraph{Limitations.}
The concept story corpus was generated by a large language model,
raising concerns about circularity in a study that probes language
model representations.
We note, however, that the generating model (Claude, Anthropic) and
the probed models (Gemma, Qwen, Llama) come from different
organisations and architectures, so the warmth/competence contrasts
are not trivially reproduced by a shared generator.
Nevertheless, the possibility that AI-generated stimuli share stylistic
properties that inflate probe accuracy across model families cannot be
fully excluded; replication with human-authored or template-based
stimuli is desirable.
The current corpus of 200 stories is a pilot that covers 50 topics;
the planned expansion to 400 stories will increase per-condition
sample size and topic diversity.
The remaining valence overlap between the warmth and competence
directions (cosine 0.75 for Gemma prior to denoising) means that
concept-level steering effects are not yet fully axis-specific;
the PCA denoising step is designed to address this and its effects are
pending.
The causal warmth steering effect on hiring callbacks is clean and
linear at 12B but absent at 27B, so the causal claim does not
generalise across scale without qualification; and within 12B, the
competence steering effect is non-monotonic at the $-0.25$ strength
level.
The probe-to-human validation correlations are moderate
($\rho = 0.355$/$0.381$ warmth and $\rho = 0.230$/$0.283$ competence
at 12B/27B), not strong, indicating that the model's internal probes
capture a component of human social perception rather than replicating
it faithfully.
The hiring prompt is simplified relative to real-world screening
contexts (single role, fixed qualifications), and results may differ
under richer evaluation settings or different job types.
The demographic disparity comparison against the Gallo--Hausladen
human callback meta-analysis remains pending the selection of
demographic groupings and callback datasets.
Finally, the architectural source of the elevated warmth/competence
vector cosine observed in Gemma-3 relative to Qwen3 and Llama-3.1
across all network depths remains an open question.

\paragraph{Future Work.}
The immediate next steps are the completion of valence denoising and the
hiring-level causal steering and benchmark evaluation.
Beyond these, the stimulus corpus will be expanded to 400 stories.
The SAE decomposition performed here on Gemma-3 can be extended to
Llama-3.1 using the Llama Scope sparse autoencoders \citep{lindsey2025llama},
enabling a model-family comparison at the feature level.
On the social-science side, extending the hiring stimuli beyond names to
category-level signals (age, disability, religion) would broaden the
scope of the benchmark comparison and test whether the representational
account generalises beyond race and gender.
Finally, translating these findings into practical debiasing interventions
--- and evaluating their robustness to adversarial prompting --- is a
natural next step for applied work.

{
\small

\clearpage
\bibliographystyle{apalike}  % CHANGED: matches authoryear natbib mode
\bibliography{references}
}



\clearpage
\begin{appendices}
\onecolumn
\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\thesection}{\Alph{section}}%
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand\thefigure{S.\arabic{figure}}
\renewcommand\thetable{S.\arabic{table}}
\begin{center}
    \begin{minipage}{\textwidth}
        \centering
        \fontsize{16}{20}\selectfont
        Supplementary Materials
    \end{minipage}
\end{center}
\vspace{.5cm}

\section*{Details on Methods}
\label{apx:methods}

\section*{Additional Results}

\end{appendices}
\end{document}
